We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-02-19 paper/paper.tex
referencias

The spatial distribution of satellite galaxies around our Milky Way
(MW) and the M31 galaxy provides a stringent test for structure
formation theories in a explicit cosmological context
\citep{2015ApJ...815...19P, 2016MNRAS.457.1931S}
The presence of a Vast Polar Structure of satellites around the Milky
Way, first suggested by \cite{1976RGOB..182..241K} and
\cite{1976MNRAS.174..695L} as a Magellanic Plane, has been
stablished in the last decade \citep{2005A&A...431..517K,2013MNRAS.435.1928P}. 
This structure is present in the $11$ brightest "classical" dwarf
galaxies and has been confirmed as fainter satellites have been
detected \citep{2014MNRAS.441.2124B}. %XXXXXXX  
In the case of M31, observational studies have found a planar satellite
distribution in a subset of $15$ satellites out of the 
total poulation of $27$ satellites.\citep{2013ApJ...766..120C,
2013Natur.493...62I}

The question whether this kind of planar configurations are atypical
in the observable Universe cannot be properly addressed given the
current observational limitations. %citation?
Comparisons against observations in large galaxy surveys have only
been able to constrain that $5\%$ of the Milky-Way like galaxies in
the Universe have two bright satellite galaxies as is the case of the
Milky Way and its Large and Small Magellanic Clouds.%citation 
A planar structure ouside the Local Group (LG) has only been confirmed
in Centaurus A \citep{muller2018whirling}.

To date, the atipicallity of satellite distributions can only be
addressed in simulations.
Most of the studies have been performed in the currently preferred
paradigm, which is based on a Cold Dark Matter cosmology in a expanding
Universe described by General Relativity including a Cosmological
Constant, the so-called Lambda Cold Dark Matter (LCDM) model.
The broad consensus is that these highly structured satellite distributions
are hard to find in simulations \citep{}.  
There are still disagreements on the precise degree of atypicallity
and the conditions required to simulate satellite distributions closer.
These disagreements seem to be originated on what is considered to be
a LG analog in the simulation.

>>>>

The spatial distribution of satellite galaxies around our Milky Way
(MW) and the M31 galaxy provides a stringent test for structure
formation theories in a explicit cosmological context
\citep{2015ApJ...815...19P, 2016MNRAS.457.1931S}.
The presence of a Vast Polar Structure of satellites around the Milky
Way, first suggested by \cite{1976RGOB..182..241K} and
\cite{1976MNRAS.174..695L} as a Magellanic Plane, has been
stablished in the last decade \citep{2005A&A...431..517K,2013MNRAS.435.1928P}. 
This structure is present in the $11$ brightest "classical" dwarf
galaxies and has been confirmed as fainter satellites have been
detected \citep{2014MNRAS.441.2124B}. 
In the case of M31, observational studies have found a planar satellite
distribution in a subset of $15$ satellites out of the 
total poulation of $27$ satellites.\citep{2013ApJ...766..120C,
2013Natur.493...62I}, although its $11$ brightest satellites (to make
a fair comparison against the MW) do not show any special feature.

The question whether this kind of planar configurations are atypical
on cosmological samples has not been properly addressed given the
current observational limitations. For instance, comparisons in large
galaxy surveys have only been able to constrain that $5\%$ of the
Milky-Way like galaxies in the Universe have two bright satellite
galaxies as is the case of the Milky Way and its Large and Small
Magellanic Clouds \citep{2011ApJ...733...62L}, or
that pairs of bright satellites preferentially show anti-correlated
when they are diametrically oposed \citep{2014Natur.511..563I}.
A planar structure outside the Local Group (LG) has been confirmed
in Centaurus A \citep{muller2018whirling} a galaxy that is only $\sim
4$Mpc away.

To date, the atipicallity of satellite distributions can only be
addressed in simulations.
First studies 
Most of the studies have been performed in the currently preferred
paradigm, which is based on a Cold Dark Matter cosmology in a expanding
Universe described by General Relativity including a Cosmological
Constant, the so-called Lambda Cold Dark Matter (LCDM) model.
The broad consensus is that these highly structured satellite distributions
are hard to find in simulations \citep{}.  
There are still disagreements on the precise degree of atypicallity
and the conditions required to simulate satellite distributions closer.
These disagreements seem to be originated on what is considered to be
a LG analog in the simulation.
